---
title: "Dog breeds Analysis"
author: "Dr. Shashikanth Vydyula"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: false
    highlight: pygments
    code_folding: show
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>


***

```{r steup, include=FALSE}
knitr::opts_chunk$set(
  comment = NA,
  warning= FALSE,
  message = FALSE
)
```

# Load Packages
```{r}
library('tidyverse') # essential R packages for data science
library('data.table')

# general visualization
library('ggplot2') # visualization
library('datasauRus') # visualization
library("ggtext") #plot text
library('plotly')

# general data manipulation
library('dplyr') # data manipulation
library('readr') # input/output
library('skimr') # overview
library('tibble') # data wrangling
library('tidyr') # data wrangling

# file referencing
library('here') # refer files
```

***


# Import Data
```{r}
dogs<-read.csv("C:/Shashi/Data & Dashboards/Datasets/dog_breeds.csv")
str(dogs)
head(dogs)
```

***

# Identifying most common problem
```{r}
#Lower case common health problems and splitting them with comma(,) delimiter
dogs<-dogs %>% mutate(Common.Health.Problems=tolower(Common.Health.Problems))

dogs<-separate(dogs, Common.Health.Problems, into=c('common_problem1', 'common_problem2', 'common_problem3'), sep=',')

head(dogs)

#Remove unwanted spaces from health problems columns
dogs$common_problem1<-trimws(dogs$common_problem1, which=c('both'))
dogs$common_problem2<-trimws(dogs$common_problem2, which=c('both'))
dogs$common_problem3<-trimws(dogs$common_problem3, which=c('both'))
head(dogs)

most_common_problem1<-dogs %>% 
  group_by(common_problem1) %>% 
  summarise(total_count=n(),
            .groups='drop')%>% 
  arrange(desc(total_count))%>% as.data.frame()


most_common_problem2<-dogs %>% 
  group_by(common_problem2) %>% 
  summarise(total_count=n(),
            .groups='drop') %>% 
  arrange(desc(total_count))%>% as.data.frame()



most_common_problem3<-dogs %>% 
  group_by(common_problem3) %>% 
  summarise(total_count=n(),
            .groups='drop') %>% 
  arrange(desc(total_count))%>% as.data.frame()

#Renaming problems column to problem for joining
most_common_problem1<-most_common_problem1 %>% rename('problem'='common_problem1')
most_common_problem2<-most_common_problem2 %>% rename('problem'='common_problem2')
most_common_problem3<-most_common_problem3 %>% rename('problem'='common_problem3')



common_problems<-full_join(most_common_problem1, most_common_problem2, by='problem') %>% 
  full_join(.,most_common_problem3, by='problem')

#Replacing NA with zero values
common_problems[is.na(common_problems)]<-0

#summing up common problem count from all the columns
common_problems$problem_count<-common_problems$total_count.x+common_problems$total_count.y+common_problems$total_count

head(common_problems)


common_problems<-common_problems %>% select(problem, problem_count) %>% 
  arrange(desc(problem_count))
head(common_problems)

top_10_problems<-common_problems %>% slice(1:10)
head(top_10_problems)

top_10_problems$problem  <- with(top_10_problems, reorder(problem, problem_count))
top_10_problems_plot<-ggplot(top_10_problems, aes(label=problem_count, fill=problem, y=problem_count, x=reorder(problem, desc(problem_count))))+
  geom_bar(stat='identity')+
  geom_text(size = 3, position = position_stack(vjust=0.5))+
  labs(x="Top 10 Disease", y=" ")+
  coord_flip()
top_10_problems_plot

```

***

# Which Country has most breeds
```{r}
#Which country has most dog breeds
country<-dogs %>% 
  group_by(Country.of.Origin) %>% 
  summarise(total_count=n(),
            .groups = 'drop') %>% 
  arrange(desc(total_count))
head(country)


country %>%
  arrange(desc(total_count)) %>% 
  slice(1:10) %>% 
  plot_ly(x= ~total_count, y= ~Country.of.Origin, type = 'bar', orientation = 'h',
          marker = list(color = 'rgb(158,202,225)',
                        line = list(color = 'rgb(8,48,107)', width = 1.5)),
          text = ~paste0('*',total_count), textposition = 'auto', hoverinfo = 'text') %>%
  layout(yaxis = list(categoryorder = "total ascending"),
         title = 'Top 10 Countries with most Breeds')
```

***

# Common fur color
```{r}
fur_color<-dogs %>% select(Breed, Fur.Color)
head(fur_color)

#lower case color names
fur_color<-fur_color %>% mutate(Fur.Color=tolower(Fur.Color))

#Split color names to different columns
fur_color<-separate(fur_color, Fur.Color, into=c("color1", "color2", "color3", "color4"), sep=',')

head(fur_color)

#remove leading and trailing spaces
fur_color$color1<-trimws(fur_color$color1, which=c('both'))
fur_color$color2<-trimws(fur_color$color2, which=c('both'))
fur_color$color3<-trimws(fur_color$color3, which=c('both'))
fur_color$color4<-trimws(fur_color$color4, which=c('both'))

fur_color[is.na(fur_color)]<-""

#most common color 1
most_common_color1<-fur_color %>% 
  group_by(color1) %>% 
  summarise(total_count=n(),
            .groups = 'drop') %>% 
  arrange(desc(total_count))
head(most_common_color1)

#most common color 2
most_common_color2<-fur_color %>% 
  group_by(color2) %>% 
  summarise(total_count=n(),
            .groups = 'drop') %>% 
  arrange(desc(total_count))
head(most_common_color2)

#most common color 3
most_common_color3<-fur_color %>% 
  group_by(color3) %>% 
  summarise(total_count=n(),
            .groups = 'drop') %>% 
  arrange(desc(total_count))
head(most_common_color3)

#most common color 4
most_common_color4<-fur_color %>% 
  group_by(color4) %>% 
  summarise(total_count=n(),
            .groups = 'drop') %>% 
  arrange(desc(total_count))
head(most_common_color4)

#renaming most_common_color to color
most_common_color1<-most_common_color1 %>% rename("color"="color1")
most_common_color2<-most_common_color2 %>% rename("color"="color2")
most_common_color3<-most_common_color3 %>% rename("color"="color3")

#joining all colors into one table
common_colors<-full_join(most_common_color1, most_common_color2, by='color') %>% 
  full_join(.,most_common_color3, by='color')

head(common_colors)

#replacomg NA values with 0
common_colors[is.na(common_colors)]<-0

#summing up common color values from all columns
common_colors$color_count<-common_colors$total_count.x+common_colors$total_count.y+common_colors$total_count

common_colors<-common_colors %>% select(color, color_count) %>% 
  arrange(desc(color_count)) %>% 
  slice(2:11) #removes first blank row count and select top
head(common_colors)

common_colors_plot<-
  common_colors %>%
  arrange(desc(color_count)) %>% 
  slice(1:10) %>% 
  plot_ly(x= ~color_count, y= ~color, type = 'bar', orientation = 'h',
          marker = list(color = 'rgb(	50, 171, 96)',
                        line = list(color = 'rgb(	120,0,128)', width = 1.5)),
          text = ~paste0('*',color_count), textposition = 'auto', hoverinfo = 'text') %>%
  layout(yaxis = list(categoryorder = "total ascending"),
         title = 'Top 10 Colors')
common_colors_plot
```
***

# Age vs Height

```{r,echo=FALSE, out.width="85%", out.height="100%", warning=FALSE}
dogs$mean_Height..in. <- sapply(strsplit(as.character(dogs$Height..in.) , "-", fixed = TRUE), function(x) sum(as.numeric(x))) 
dogs$mean_Height..in. <- dogs$mean_Height..in. / 2 
dogs$mean_Longevity..yrs. <- sapply(strsplit(as.character(dogs$Longevity..yrs.) , "-", fixed = TRUE), function(x) sum(as.numeric(x))) 
dogs$mean_Longevity..yrs. <- dogs$mean_Longevity..yrs. / 2 

#Scatter plot with all breed names
plot(dogs$mean_Longevity..yrs., dogs$mean_Height..in.)+
  text(dogs$mean_Longevity..yrs., dogs$mean_Height..in.,dogs$Breed)
```

***


```{r}
#Bar plot (Age vs Height)
height_vs_age_plot<-ggplot(dogs, aes(fill=Breed, y=mean_Height..in., x=mean_Longevity..yrs.))+
  geom_bar(position = 'dodge', stat='identity')+
  labs(x="Longevity in Years", y="Height in inches")+
  theme(legend.position="none")

height_vs_age_plot

#Mean age 
mean_age <- ggplot(dogs, aes(x=mean_Longevity..yrs.))

mean_age + geom_area(stat = "bin", fill = "lightblue")+
  geom_vline(aes(xintercept=mean(mean_Longevity..yrs.)),
             color="blue", linetype="dashed", size=1)+
  labs(x="Mean Age of Dogs", y="")

#Correlating mean age with top-5 countries with most breeds
p<-ggplot(dogs[dogs$Country.of.Origin %in% c("England","Germany","Australia", "France", "Scotland"),], aes(x=mean_Longevity..yrs., fill=Country.of.Origin))+
  geom_area(stat ="bin")+facet_grid(Country.of.Origin ~ .)+
  labs(x="Mean Age of Dogs", y="")
p

#joining age vs height with mean age of top-5 countries with most breeds
ggplot(dogs[dogs$Country.of.Origin %in% c("England","Germany","Australia", "France", "Scotland"),], aes(x=mean_Longevity..yrs., y=mean_Height..in.)) +
  geom_area( fill="#69b3a2", alpha=0.4) +
  geom_line(color="#69b3a2", size=2) +
  geom_point(size=3, color="#69b3a2") +
  ggtitle("Age vs Height")+
  labs(x="Mean Longevity", y="Mean height")+
  geom_vline(aes(xintercept=mean(mean_Longevity..yrs.)),
             color="black", linetype="dashed", size=1)


```

***

# Conclusion

**Dental issues is the most common health problem.**


**Black is the most common fur color.**


**England has more breeds (24) than other countries.**


**Mean age of dogs is 12.5 years but most of the dogs have average life of 12.5 to 14 years**


**Larger dogs have lesser life span compared to smaller dogs**