---
title: "Population Growth Prediction"
author: "Dr. Shashikanth Vydyula"
date: "`r Sys.Date()`"
output: 
  html_document:
    df_print: paged
    toc: true
    number_sections: false
    highlight: pygments
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = NA,
  warning= FALSE,
  message = FALSE,
  fig.align='center'
   
    
)

```


```{r}
required_packages<-c("tidyverse", "data.table", "dplyr", "readr", "skimr", "tibble", "tidyr", "stringr", "forcats","janitor", "Tmisc", "class", "lubridate", "forecast", "timetk", "zoo", "highcharter", "gganimate", "geonames", "ggmap", "maps", "plotly", "RColorBrewer" , "htmltools", "htmlwidgets", "tm", "gridExtra", "earth")

invisible(lapply(required_packages, require, character.only = TRUE, quietly = TRUE))

```

# Data Load and Manipulation
```{r}
population<-fread("C:/Shashi/Data & Dashboards/Datasets/Population Growth.csv")

str(population)

population$`Population Growth Rate`<-as.numeric(gsub(",","", population$`Population Growth Rate`))
population$`Growth Rate`<-as.numeric(gsub("%", "", population$`Growth Rate`))

#accumulate function
accumulate_by = function(dat, var) {
  var <- lazyeval::f_eval(var, dat)
  lvls <- plotly:::getLevels(var)
  dats <- lapply(seq_along(lvls), function(x) {
    cbind(dat[var %in% lvls[seq(1, x)], ], frame = lvls[[x]])
  })
  dplyr::bind_rows(dats)
}

df <- population %>%
  accumulate_by(~Year)
str(population)
```

# Population Growth
```{r}
highchart() %>% 
  hc_chart(type="spline", options3d=list(enabled=T, alpha=10, beta=15)) %>% 
  hc_xAxis(categories=population$Year) %>% 
  hc_add_series(data=population$`Population Growth Rate`, name="Population Growth Rate") %>% 
  hc_add_theme(hc_theme_bloom()) %>% 
  hc_title(text="Population Growth (1950-2023)")

plot_ly() %>%
  add_trace(
    x = ~Year, 
    y = ~`Population Growth Rate`,
    name="Population",
    frame = ~frame,
    type = 'scatter',
    mode = 'lines',
    data = df,
    opacity = 1.0) %>% 
  layout(
    title = "Population Growth (in billions)"
  ) %>%
  animation_opts(
    frame = 85, 
    transition = 0, 
    redraw = F
  ) %>%
  animation_slider(
    hide = T
  ) %>%
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  ) %>% 
  layout(xaxis = list(title = "Year"), yaxis = list(title = "Population"))
```

# Growth Rate %

```{r}
highchart() %>%
  hc_chart(type = "line") %>%
  hc_xAxis(categories=population$Year) %>% 
  hc_add_series(data=population$`Growth Rate`, name="Growth Rate") %>%
  hc_tooltip(valueDecimals = 2, valueSuffix = "%") %>% 
  hc_add_theme(hc_theme_sandsignika()) %>% 
  hc_title(text="Growth Rate % (1950-2023)")

plot_ly() %>%
  add_trace(
    x = ~Year, 
    y = ~`Growth Rate`,
    name="Population",
    frame = ~frame,
    type = 'scatter',
    mode = 'lines',
    data = df,
    opacity = 1.0) %>% 
  layout(
    title = "Growth Rate %"
  ) %>%
  animation_opts(
    frame = 85, 
    transition = 0, 
    redraw = F
  ) %>%
  animation_slider(
    hide = T
  ) %>%
  animation_button(
    x = 1, xanchor = "right", y = 0, yanchor = "bottom"
  ) %>% 
  layout(xaxis = list(title = "Year"), yaxis = list(title = "Rate"))

```


# Prediction Models (Next 25 years: 2024-2048)

```{r}
dataset<-fread("C:/Shashi/Data & Dashboards/Datasets/Population Growth.csv")


colnames(dataset)[2]<-"population.growth"
colnames(dataset)[3]<-"growth.percent"
colnames(dataset)<-tolower(colnames(dataset))

dataset$population.growth <- as.numeric(gsub(",", "", dataset$population.growth))
dataset$growth.percent <- as.numeric(gsub("%", "", dataset$growth.percent))
```

## Model 1- Linear Regression
```{r}
model_lm <- lm(population.growth ~ year, data = dataset)

new_years <- 2024:2048
new_data <- data.frame(year = new_years)
predicted_population_lm <- predict(model_lm, newdata = new_data)

initial_population_lm <- dataset$population.growth[length(dataset$population.growth)]
predicted_growth_rate_lm <- ((predicted_population_lm - initial_population_lm) / initial_population_lm) * 100
predicted_growth_rate_lm <- paste0(round(predicted_growth_rate_lm, 2), "%")

lm_result <- data.frame(
  Year = new_years,
  Population_Growth_Rate = predicted_population_lm,
  Growth_Rate = predicted_growth_rate_lm
)

print(lm_result)

l1<-highchart() %>% 
  hc_chart(type="spline") %>% 
  hc_xAxis(categories=lm_result$Year) %>% 
  hc_add_series(data=lm_result$Population_Growth_Rate, name="Population Growth (in billions)") %>% 
  hc_title(text="Linear Regression Prediction - Population Growth (in billions)") %>% 
  hc_add_theme(hc_theme_google())

lm_result$Growth_Rate<-as.numeric(gsub("%", "", lm_result$Growth_Rate))

l2<-highchart() %>% 
  hc_chart(type="spline") %>% 
  hc_xAxis(categories=lm_result$Year) %>% 
  hc_add_series(data=lm_result$Growth_Rate, name="Growth Rate (%)") %>% 
  hc_title(text="Linear Regression Prediction - Growth Rate(%)") %>% 
  hc_tooltip(valueDecimals = 2, valueSuffix = "%") %>% 
  hc_add_theme(hc_theme_google())

highcharter::hw_grid(
  l1,
  l2) %>% 
  htmltools::browsable()
```

## Model 2 - ARIMA (AutoRegressive Integrated Moving Average) Prediction
```{r}
ts_data <- ts(dataset$population.growth, start = 1950, end = 2023, frequency = 1)

model_arima <- auto.arima(ts_data)

forecast_result <- forecast(model_arima, h = 25)

initial_population_arima <- dataset$population.growth[length(dataset$population.growth)]
predicted_growth_rate_arima <- ((forecast_result$mean - initial_population_arima) / initial_population_arima) * 100
predicted_growth_rate_arima <- paste0(round(predicted_growth_rate_arima, 2), "%")

arima_result <- data.frame(
  Year = 2024:2048,
  Population_Growth_Rate = forecast_result$mean,
  Growth_Rate = predicted_growth_rate_arima
)
print(arima_result)

a1<-ggplot(data = arima_result, aes(x = Year, y = Population_Growth_Rate)) +
  geom_line() +
  labs(x = "Year", y = "Population Growth (in billions)") +
  ggtitle("ARIMA Model Prediction - Population Growth (billions)") +
  theme_minimal() +  # Apply minimal theme
  theme(
    plot.title = element_text(color = "darkblue", size = 16, face = "bold"),
    axis.title.x = element_text(color = "darkgreen", size = 14),
    axis.title.y = element_text(color = "darkgreen", size = 14),
    axis.text = element_text(color = "black", size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()
  )

a2<-ggplot(data = arima_result, aes(x = Year, y = Growth_Rate, group = 1)) +
  geom_line() +
  labs(x = "Year", y = "Population Growth Rate (%)") +
  ggtitle("ARIMA Model Prediction - Population Growth Rate (%)") +
  theme_minimal() +
  theme(
    plot.title = element_text(color = "darkblue", size = 16, face = "bold"),
    axis.title.x = element_text(color = "darkgreen", size = 14),
    axis.title.y = element_text(color = "darkgreen", size = 14),
    axis.text = element_text(color = "black", size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()
  )

a1

a2

```

## Model 3 - MARS (Multivariate Adaptive Regression Splines) Prediction 
```{r}
model_data<-fread("C:/Shashi/Data & Dashboards/Datasets/Population Growth.csv")

colnames(model_data)[1]<-"years"
colnames(model_data)[2]<-"population_growth"
colnames(model_data)[3]<-"growth_rate"

model_data$population_growth <- as.numeric(gsub(",", "", model_data$population_growth))
model_data$growth_rate <- as.numeric(gsub("%", "", model_data$growth_rate))



model_mars_growth <- earth(population_growth ~ years, data = model_data)

new_years <- 2024:2048
new_data <- data.frame(Year = new_years)
predicted_population_growth <- predict(model_mars_growth, newdata = new_data)

model_mars_rate <- earth(growth_rate ~ years, data = model_data)

predicted_growth_rate <- predict(model_mars_rate, newdata = new_data)

mars_result <- data.frame(
  Year = new_years,
  Population_Growth = predicted_population_growth,
  Growth_Rate = predicted_growth_rate
)

m1<-highchart() %>% 
  hc_chart(type="spline") %>% 
  hc_xAxis(categories=mars_result$Year) %>% 
  hc_add_series(data=mars_result$population_growth, name="Population Growth (in billions)") %>% 
  hc_title(text="MARS Prediction - Population Growth") %>% 
  hc_add_theme(hc_theme_google())

m2<-highchart() %>% 
  hc_chart(type="spline") %>% 
  hc_xAxis(categories=mars_result$Year) %>% 
  hc_add_series(data=mars_result$growth_rate, name="Growth Rate (%)") %>% 
  hc_title(text="MARS Prediction - Growth Rate") %>% 
  hc_add_theme(hc_theme_google())


highcharter::hw_grid(
  m1,
  m2) %>% 
  htmltools::browsable()


```

<div style="color:white;
           display:fill;
           border-radius:5px;
           background-color:#5642C5;
           font-size:110%;
           font-family:Verdana;
           letter-spacing:0.5px">

<p style="padding: 10px;
              color:white;">
Linear Regression and ARIMA model predict positive growth rate of population in next 25 years. Whereas MARS model predicts a decrease in growth rate. This can be due to the ability of MARS model to capture nonlinear relationships and interactions between variables. It uses piece wise linear segments to model the data, which allows it to fit more complex patterns. Linear regression and ARIMA, being linear models, have less flexibility in capturing nonlinear patterns.
</p>
</div>


***

<div style="color:white;
           display:fill;
           border-radius:5px;
           background-color:#5642C5;
           font-size:110%;
           font-family:Verdana;
           letter-spacing:0.5px">

<p style="padding: 10px;
              color:white;">
As we see in the Exploratory data analysis of this data set, there is a decrease in population growth rate since past 30 years (except in 2023 where there is an increase compared to 2022). As per this trend, MARS prediction might have interpreted that there will be a negative growth rate in future.
</p>
</div>