---
title: "World Alcohol Consumption"
author: "Dr. Shashikanth Vydyula"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: false
    highlight: pygments
    code_folding: show
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>


***

```{r steup, include=FALSE}
knitr::opts_chunk$set(
  comment = NA,
  warning= FALSE,
  message = FALSE
)
```
***

# Install Packages
```{r}
library('tidyverse') 
library('readr')
library('lubridate')
library('ggplot2')
library('tidyr')
library('dplyr')

```

***

# Import Dataset
```{r}
alcolohol_consumption<-read.csv("C:/Shashi/Data & Dashboards/Datasets/World Alcoholic Consumption datasets.csv")

head(alcolohol_consumption)
summary(alcolohol_consumption)
```

***

# Data Manipulation
```{r}
alcolohol_consumption$date <- lubridate::as_date(mdy(alcolohol_consumption$Date)) #Formatting Date column


str(alcolohol_consumption)
summary(alcolohol_consumption)
```
***
```{r}
# Gender count

gender_count<-alcolohol_consumption %>% count(Gender)
gender_count
```


**Women consumers are more than men**

***
```{r}
# Country count
country_count<-alcolohol_consumption %>% 
  count(Countries)
head(country_count)
```
***

# Top alcoholic countries
```{r}
top_countries<-country_count[order(country_count$n, decreasing = TRUE),]
view(top_countries)
top_10_countries<-top_countries %>% slice(1:10)

top_10_countries$Countries  <- with(top_10_countries, reorder(Countries, n))
top_10_countries_plot<-ggplot(top_10_countries,aes(label=n,fill=Countries, y=n, x=reorder(Countries, desc(n))))+
  geom_bar(stat="identity")+
  geom_text(size = 3, position = position_stack(vjust = 0.5))+
  labs(x="Country", y="Count")
top_10_countries_plot
```


**China Ranks first**

***


```{r}
mapdata <- map_data("world")
head(mapdata)

top_countries<- top_countries %>% rename("region"="Countries")

top_countries$region <- recode(top_countries$region
                                   ,'United States' = 'USA'
                                   ,'United Kingdom' = 'UK'
                                   , 'United Arab Emirates' ='UAE'
)




mapdata <- 
  left_join(mapdata, top_countries, by='region')

theme1 <- theme(
  axis.text = element_blank(),
  axis.line = element_blank(),
  axis.ticks = element_blank(),
  panel.border = element_blank(),
  panel.grid = element_blank(),
  axis.title = element_blank(),
  panel.background = element_rect(fill = "grey85"),
  plot.title = element_text(hjust = 0.5)
)

ggplot(mapdata, aes(x=long, y=lat, group=group)) + 
  coord_fixed(1.3) +
  geom_polygon(aes(fill = n)) +
  scale_fill_distiller(palette ="RdBu", direction = -1) + 
  ggtitle("World Alcohol Consumption") +
  theme1 




```

***

# Top alcoholic countries with male and female count
```{r}
#Male and Female count
male_female_count<-alcolohol_consumption %>% 
  group_by(Countries, Gender) %>% 
  summarise(total_count=n(),.groups = 'drop')

view(male_female_count)

top_10_male_female_count<-
  male_female_count[order(male_female_count$total_count, decreasing = TRUE),] %>% 
  slice(1:19)

head(top_10_male_female_count)

top_10_male_female_count$Countries<-with(top_10_male_female_count,reorder(Countries,total_count))

top_10_m_f_plot<-
  ggplot(top_10_male_female_count, aes(label=total_count, fill=Gender, y=total_count,x=Countries))+
  geom_bar(stat="identity")+
  geom_text(size=3, position=position_stack(vjust = 0.5))+
  labs(X="Country", y="Gender Count")+
  coord_flip()
top_10_m_f_plot

```

***



# Year with highest alcohol consumptionconsumption
```{r}
##Before calculating the year with highest consumption we need to separate year from Date
alcolohol_consumption$year <- lubridate::year(alcolohol_consumption$date)


#let us separate month and date as well
alcolohol_consumption$month <- lubridate::month(alcolohol_consumption$date) #separates month number
alcolohol_consumption$month_name<-month.name[alcolohol_consumption$month] #gives name to the month by its number

head(alcolohol_consumption)

#Top Year (total)
top_year_total<-alcolohol_consumption %>% 
  group_by(year, Gender) %>% 
  summarise(total_count=n(),.groups = 'drop')

view(top_year_total)

top_year_total_plot<-ggplot(top_year_total, aes(label=total_count, fill=Gender, y=total_count, x=year))+
  geom_bar(stat="identity")+
  geom_text(size=4, position=position_stack(vjust = 0.5))+
  labs(x="Year", y="Count by year and Gender")

top_year_total_plot

```


**2020 recorded high alcohol consumption than 2019.**


***

# Month with highest alcohol consumption
```{r}
top_month_total<-alcolohol_consumption %>% 
  group_by(month_name, Gender) %>% 
  summarise(total_count=n(),.groups = 'drop')
view(top_month_total)


top_month_total_plot<-ggplot(top_month_total, aes(label=total_count, fill=Gender, y=total_count, x=reorder(month_name, total_count)))+
  geom_bar(stat="identity")+
  geom_text(size=3, position=position_stack(vjust=0.25))+
  facet_grid(~Gender)+
  labs(x="Month in Descending Order of alcohol Consumption", y="Total Count")+
  coord_flip()

top_month_total_plot
```


**May has highest alcohol consumption and December has lowest.**

***

# Which month women consumed most
```{r}
women_only <- top_month_total[ which(top_month_total$Gender=='Female'), ]

view(women_only)

women_only_plot<-ggplot(women_only, aes(label=total_count, fill=month_name, y=total_count, x=reorder(month_name, total_count)))+
  geom_bar(stat="identity")+
  geom_text(size=3, position=position_stack(vjust=0.25))+
  facet_grid(~Gender)+
  labs(x="Month", y="Count")+
  coord_flip()+
  theme(legend.position = "none")

women_only_plot
```


**Women consumed more alcohol in February.**

***

# Which month men consumed most
```{r}
men_only<-top_month_total[which(top_month_total$Gender=='Male'),]
view(men_only)

men_only_plot<-ggplot(men_only, aes(label=total_count, fill=month_name, y=total_count, x=reorder(month_name, total_count)))+
  geom_bar(stat="identity")+
  geom_text(size=3, position=position_stack(vjust=0.25))+
  facet_grid(~Gender)+
  labs(x="Month", y="Count")+
  coord_flip()+
  theme(legend.position = "none")

men_only_plot
```


**Men consumed more alcohol in May.**

***




<div style="color:white;
           display:fill;
           border-radius:5px;
           background-color:#5642C5;
           font-size:110%;
           font-family:Verdana;
           letter-spacing:0.5px">

<p style="padding: 25px;
              color:white;
              text-align: center">
              Thank you🤗
</p>
</div>

***
