---
title: "World’s Best Employers TOP 100"
author: "Dr. Shashikanth Vydyula"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    number_sections: false
    highlight: pygments
    code_folding: show
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: DarkRed;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkRed;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
  text-align: center;
}
</style>


***

```{r steup, include=FALSE}
knitr::opts_chunk$set(
  comment = NA,
  warning= FALSE,
  message = FALSE
)
```
***

```{r}
# Package for data science
library('tidyverse') # essential R packages for data science
library('data.table')

# general visualization
library('ggplot2') # visualization
library('datasauRus') # visualization

# general data manipulation
library('dplyr') # data manipulation
library('readr') # input/output
library('skimr') # overview
library('tibble') # data wrangling
library('tidyr') # data wrangling
library('purrr') # data wrangling
library('stringr') # string manipulation
library('forcats') # factor manipulation
library('janitor') # data cleaning
library('Tmisc') # data manipulation
library('class') # functions for classification

# Date + forecast
library('lubridate') # date and time
library('forecast') # time series analysis
library('timetk') # time series analysis

# file referencing
library('here') # refer files

#others
library('shiny')
library('leaflet')
library('ggmap')
library('plotly')
library('countrycode')
library('mice')
library('GGally')
library('sf')


```


# Import Data set
```{r}
best_employees<-read.csv("C:/Shashi/Data & Dashboards/Datasets/Worlds Best Employers.csv")

head(best_employees)
summary(best_employees)
```

**The employees column is formatted as character with commas in between. I will format it to numeric.**
```{r}

best_employees$EMPLOYEES<-as.numeric(gsub(",","", best_employees$EMPLOYEES))
summary(best_employees)

```
# Top 10 Companies

```{r}
top_10_companies<-best_employees %>% slice(1:10)

tibble(top_10_companies)
```
**Samsung Electronics ranks 1st in the world**

***

# Company with highest number of employees

```{r}
highest_numer_of_employees<-
  best_employees[order(best_employees$EMPLOYEES, decreasing = TRUE),] %>% 
  slice(1:10)
view(highest_numer_of_employees)

highest_numer_of_employees$NAME  <- with(highest_numer_of_employees, reorder(NAME, EMPLOYEES))

highest_employees<-ggplot(highest_numer_of_employees, aes(fill=NAME, y=EMPLOYEES, x=reorder(NAME, -EMPLOYEES)))+
  geom_bar(position="dodge", stat="identity") +
             labs(x="Company", y="Number of Employees")+
  coord_flip()

highest_employees
```


**Amazon has the Highest number of employees**

***

# Which industry has most companies
```{r}
top_10_industry<-best_employees %>% count(INDUSTRIES)

top_10_industry<-
  top_10_industry[order(top_10_industry$n, decreasing = TRUE),] %>% 
  slice(1:10)
view(top_10_industry)

plot_ly(data=top_10_industry,labels=~INDUSTRIES, values=~n, type="pie") %>% 
  layout(autosize = F, width = 1080, height = 450,title = 'Top 10 Industries',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = TRUE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```

> **IT, Internet, Software & services industry has highest companies (12) followed by Aerospace & Defense, Automotive, and Banking & Financial Services with 10 respective companies each.**

***

# Which country has highest number of companies
```{r}
country_with_highest_companies<-best_employees %>% 
  count(COUNTRY.TERRITORY)

country_with_highest_companies<-
  country_with_highest_companies[order(country_with_highest_companies$n, decreasing = TRUE),] %>% 
  slice(1:10)

view(country_with_highest_companies)

fig <- plot_ly(country_with_highest_companies, x = ~COUNTRY.TERRITORY, y = ~n, text = ~COUNTRY.TERRITORY, type = 'scatter', mode = 'markers',
               marker = list(size = ~n, opacity = 0.5, color = 'rgb(255, 54, 54)'))
fig <- fig %>% layout(title = 'Country with highest number of companies',
                      xaxis = list(showgrid = FALSE),
                      yaxis = list(showgrid = FALSE))
fig

```
**USA has 40 top 100 companies**

***

# Company with least number of employees
```{r}
least_employees<-best_employees[order(best_employees$EMPLOYEES, decreasing = FALSE),] %>% 
  slice(1:10)
view(least_employees)

least_employees$NAME  <- 
  with(least_employees, reorder(NAME, EMPLOYEES))

least_employees_plot<-ggplot(least_employees, aes(fill=NAME, y=EMPLOYEES, x=reorder(NAME, -EMPLOYEES)))+
  geom_bar(position="dodge", stat="identity") +
  labs(x="Company", y="Number of Employees") +
  coord_flip()
  
ggplotly(least_employees_plot)


```

> **Mertiz Financial Group has least number of employees (20).**

# Which industry has most employees in top 100
```{r}
top_10_industry_employees<-best_employees %>% 
  count(INDUSTRIES, EMPLOYEES)

top_10_industry_employees<- top_10_industry_employees %>% count(INDUSTRIES, wt=EMPLOYEES)

top_10_industry_employees<-top_10_industry_employees[order(top_10_industry_employees$n, decreasing = TRUE),] %>% 
  slice(1:10)

view(top_10_industry_employees)

plot_ly(data=top_10_industry_employees,labels=~INDUSTRIES, values=~n, type="pie") %>% 
  layout(autosize = F, width = 1080, height = 450,title = 'Top 10 Industries with highest number of employees',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = TRUE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))


```
**IT, Internet, Software & services has highest number with over 2.6 million employees**

***

<div class="alert alert-block alert-info">
**CONCULSION:**


1. Samsung Ranks first in best company of the world.


2. Amazon has highest number of employees with nearly 1.6 million.


3. IT, Internet, Software & Services industry has highest number of employees with over 2.6 million.


4. Mertiz Financial Group has least number of employees (20).


5. USA has 40 companies that rank in top 100 World Best Employers.


</div>
